---
title: "Lesson 5 - Exploring Two Variables"
output: html_notebook
---

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep='\t') 
ggplot(aes(x = age, y = friend_count), data=pf) +
  geom_point() + xlim(13, 90)
```

***

#### What are some things that you notice right away?
Response:
- People under the age of age have far more friends that people over 30.
- People over age 90 have surprisingly more friends.
***


### Overplotting
Notes:
- SOme points are spread out.
- Some are stacked on top of each other.
- Use alpha parameter 1/20 -> It will take 20 points to be equivalent of one data point.
- Age is a continuous variable, but in the graph it appears to be.. To add the effect of continuity, use jitter to add noise. Observe the difference.
```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) +
  xlim(13, 90)

ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20) +
  xlim(13, 90)
  
```

#### What do you notice in the plot?
Response:
- The bulk of users have really have friend counts less than 1000.

***

### Coord_trans()
Notes:
- To get a better visualisation of the data, limit y axis

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) +
  xlim(13, 90) +
  coord_trans(y = 'sqrt')
```

#### What do you notice?
- Easier to see the friend count.
- Cant use jitter directly. Some people have 0 friend count. They would have imaginary values after sqrt application. TO add jitter use geom_point(position = position_jitter())
```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13, 90) +
  coord_trans(y = 'sqrt')
```


***

### Alpha and Jitter
Notes:
- Explore relationship between friendship initiated and age
```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 1/10, position = position_jitter(h = 0)) +
  xlim(13, 90) +
  coord_trans(y = 'sqrt')
```

***

### Overplotting and Domain Knowledge
Notes:
- Make adjustments to scatter plot even further !!
- Transform axis into percentages to better understand data.
- 
***

### Conditional Means
Notes:
- Summarize relationship between 2 variables.
- Scatterplot doesnt really help us understand the intricacies.
- 'dplyr' helps split the dataframe and apply functions to it.

```{r Conditional Means}
install.packages('dplyr')
library(dplyr)

# How does average frindcount vary over  age?
age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())
head(pf.fc_by_age)

# Alternate COde : Chaining
pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n())
```

Create your plot!
Average friend_count vs Age
```{r Conditional Means Plot}
ggplot(pf.fc_by_age, aes(x=age, y=friend_count_mean)) + geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(pf, aes(x=age, y=friend_count)) + xlab('Age') + ylab('Friend Count') + 
  geom_point(alpha=1/20, position = position_jitter(h=0), color='orange') + 
  xlim(13,90) + 
  coord_trans(y='sqrt') + 
  geom_line(stat='summary', fun.y=mean) + 
  geom_line(stat ='summary', fun.y=quantile, fun.args=list(probs=0.1), linetype=2, color='blue') + 
  geom_line(stat ='summary', fun.y=quantile, fun.args=list(probs=0.9), linetype=2, color='blue') +
  geom_line(stat ='summary', fun.y=median, color='blue')

ggplot(pf, aes(x=age, y=friend_count)) + xlab('Age') + ylab('Friend Count') + 
  geom_point(alpha=0.05, position = position_jitter(h=0), color='orange') +  
  coord_cartesian(xlim = c(13,70), ylim = c(0,1000)) + 
  geom_line(stat='summary', fun.y=mean) +
  geom_line(stat ='summary', fun.y=quantile, fun.args=list(probs=0.1), linetype=2, color='blue') + 
  geom_line(stat ='summary', fun.y=quantile, fun.args=list(probs=0.9), linetype=2, color='blue') +
  geom_line(stat ='summary', fun.y=median, color='blue')
```

#### What are some of your observations of the plot?
Response:
- Having more than 1000 frineds is rare.
- For 35-50 yr olds, friend_count falls below 50
- 

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:
- Summarise the strength of ther elationship using a single number.
- Use pearson_product relation coefficient
- Rule of thumb:
  -  |correlation| > 0.3 is meaningful but small
  -  |correlation| > 0.5 is moderate
  -  |correlation| > 0.7 is pretty large
```{r Correlation}
?cor.test
cor.test(pf$age, pf$friend_count, method='pearson')
# Alternate code
with(pf, cor.test(age, friend_count, method='pearson'))

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age<=70), cor.test(age, friend_count)) # default method is pearson
```

***

### Correlation Methods
Notes:
```{r}
with(subset(pf, age<=70), cor.test(age, friend_count, method='spearman') )
```

***

## Create Scatterplots
Notes:
Create a scatterplot of likes_received (y) vs. www_likes_received (x).
- USe 95th percentile using quantile command
- use lm - linear model

```{r}
ggplot( data = pf, aes(x=pf$www_likes_received, y=pf$likes_received)) +
  geom_point(alpha = .1, position = position_jitter(h=0)) +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +  
  geom_smooth(method='lm', color='red') 


```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
cor.test(pf$www_likes_received, pf$likes_received)
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)

data(Mitchell)
?Mitchell
```

Create your plot!

```{r Temp vs Month}
ggplot(Mitchell, aes(x = Mitchell$Temp, y = Mitchell$Month)) +
  geom_point(alpha = 0.5, position = position_jitter(h=0))
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:
- Months are discrete. 
```{r Making Sense of Data}
range(Mitchell$Month)
ggplot(Mitchell, aes(x=Month, y=Temp)) + 
  geom_point() +
  scale_x_discrete(breaks=seq(0,203,12))
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:
```{r}
ggplot(aes(x=(Month%%12),y=Temp), data=Mitchell)+
  geom_point()

x <- seq(0, 4*pi, pi/20)
y <- cos(x)
qplot(x = x, y = y)
dcor.ttest(x, y)
```

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
names(pf)
head(pf)
pf$age_with_months <- pf$age + (12 - pf$dob_month)/12
pf$age_with_months
```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}
pf.fc_by_age_months <- group_by(pf, age_with_months) %>%
  summarise( friend_count_mean=mean(friend_count),
             friend_count_median=median(friend_count),
             n=n()) %>%
  arrange(age_with_months)
pf.fc_by_age_months
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(subset(pf.fc_by_age_months, age_with_months<71), aes(x=age_with_months, y=friend_count_mean)) + geom_line()

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(subset(pf.fc_by_age, age<71), aes(x=age, y=friend_count_mean)) + geom_line() + geom_smooth()

p2 <- ggplot(subset(pf.fc_by_age_months, age_with_months<71), aes(x=age_with_months, y=friend_count_mean)) + geom_line() + geom_smooth()

p3 <- ggplot(subset(pf.fc_by_age, age<71), aes(x=round(age/5)*5, y=friend_count_mean)) + geom_line(stat='summary', fun.y=mean)

library(gridExtra)
grid.arrange(p2,p1,p3,ncol=1)

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***