Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)
library(dplyr)
pf <- read.csv('pseudo_facebook.tsv', sep='\t')

ggplot(aes(x=gender, y=age), data = subset(pf, !is.na(gender))) +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)


ggplot(aes(x = friend_count), data = subset(pf, !is.na(gender))) + 
  geom_histogram()


ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```
```{r}
# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

# Here is an example of the structure of your data frame. Your
# data values will be different. Note that if you are grouping by
# more than one variable, you will probably need to call the
# ungroup() function. 

#   age gender mean_friend_count median_friend_count    n
# 1  13 female          247.2953                 150  207
# 2  13   male          184.2342                  61  265
# 3  14 female          329.1938                 245  834
# 4  14   male          157.1204                  88 1201


pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n())%>%
            ungroup()%>%
            arrange(age)
pf.fc_by_age_gender
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
# Create a line graph showing the
# median friend count over the ages
# for each gender. Be sure to use
# the data frame you just created,
# pf.fc_by_age_gender.

ggplot(pf.fc__by_age_gender, aes(x = age, y = median_friend_count, color = gender)) +
  geom_line()
```

***

### Thinking in Ratios
Notes:
How many friends does a female user has than a male user?
  - Current data is in long format.
  - Long Format -> data for each year is repeated. 
  - We need to convert it into wide format.
  
```{r}
head(pf.fc_by_age_gender)
```

***

### Wide and Long Format
Notes:
- Use reshape2 package
  
***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc__by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')
head(pf.fc_by_age_gender.wide)

# We could also create a similar data frame using the dplyr and tidyr packages:
install.packages('tidyr')
library(tidyr)
pf.fc_by_age_gender.wide <-
  subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
         !is.na(gender)) %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = male / female)

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
# Plot the ratio of the female to male median
# friend counts using the data frame
# pf.fc_by_age_gender.wide.

# Think about what geom you should use.
# Add a horizontal line to the plot with
# a y intercept of 1, which will be the
# base line.

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

ggplot(pf.fc_by_age_gender.wide, aes(x = age, y = female/male)) +
  geom_line() +
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 6)
  xlab('Age') +
  ylab('Median Female Friend Count / Median Male Friend Count')
```

***

### Third Quantitative Variable
Notes:
  - Use bends
  - Explore all 4 variables - friend_count, age, gender, tenure is using a 2-D display like a scatter 
  plot. We can bend one of the quantitative variables and compare those bends.
  - Here we can group users by the year they joined
```{r Third Quantitative Variable}
# Create a variable called year_joined
# in the pf data frame using the variable
# tenure and 2014 as the reference year.


pf$year_joined <- 2014 - ceiling(pf$tenure/365)
# ALternate COde
pf$year_joined <- floor(2014 - pf$tenure/365)
head(pf$year_joined)
```

***

### Cut a Variable
Notes:
  - we can group some of the years together
  - Cut function is useful for making discrete variables from continuous/numericacl one.
  - year_joined.bucket is a categorical variable which contains bins

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)

# Create a new variable in the data frame
# called year_joined.bucket by using
# the cut function on the variable year_joined.

# You need to create the following buckets for the
# new variable, year_joined.bucket

#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

pf$year_joined.bucket <- cut(pf$year_joined, 
                             breaks=c(2004,2009,2011,2012,2014), 
                             right=TRUE)
head(pf$year_joined.bucket)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
# Create a line graph of friend_count vs. age
# so that each year_joined.bucket is a line
# tracking the median user friend_count across
# age. This means you should have four different
# lines on your plot.

# You should subset the data to exclude the users
# whose year_joined.bucket is NA.
?table
table(pf$year_joined.bucket, useNA='ifany')

ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=median)

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
# Write code to do the following:

# (1) Add another geom_line to code below
# to plot the grand mean of the friend count vs age.

# (2) Exclude any users whose year_joined.bucket is NA.

# (3) Use a different line type for the grand mean.

ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=median) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
summary(with(subset(pf, tenure>0), friend_count/tenure))
# Alternate Code
with( subset(pf, tenure >= 1), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
# Create a line graph of mean of friendships_initiated per day (of tenure)
# vs. tenure colored by year_joined.bucket.

# You need to make use of the variables tenure,
# friendships_initiated, and year_joined.bucket.

# You also need to subset the data to only consider user with at least
# one day of tenure.

ggplot(subset(pf, tenure>0), aes(x=tenure, y=friendships_initiated/tenure, color=year_joined.bucket)) + 
  geom_line(stat='summary', fun.y=mean)

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

library(gridExtra)

p1 <- ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

p2 <- ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p3 <- ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p4 <- ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

grid.arrange(p1,p2,p3,p4, ncol=1)

```

```{r}
# Instead of geom_line(), use geom_smooth() to add a smoother to the plot.
# You can use the defaults for geom_smooth() but do color the line
# by year_joined.bucket

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket))
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)

# change the id from an int to a factor
yo$id <- factor(yo$id)
str(yo)

qplot(data=yo, x=price)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
# Create a new variable called all.purchases,
# which gives the total counts of yogurt for
# each observation or household.

# One way to do this is using the transform
# function. You can look up the function transform
# and run the examples of code at the bottom of the
# documentation to figure out what it does.

# The transform function produces a data frame
# so if you use it then save the result to 'yo'!

# OR you can figure out another way to create the
# variable.

# ENTER YOUR CODE BELOW THIS LINE
# ========================================================
yo <- transform(yo, all.purchases=strawberry+blueberry+pina.colada+plain+mixed.berry)

# alternate method
yo$all.purchases <- with(yo, strawberry+blueberry+pina.colada+plain+mixed.berry)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
# Create a scatterplot of price vs time.

# This will be an example of a time series plot.

# Resolve overplotting issues by using
# techniques you learned in Lesson 4.

# What are some things that you notice?

# ENTER YOUR CODE BELOW THIS LINE
# ================================================
ggplot(yo, aes(x=time, y=price)) + 
  geom_point(alpha=0.1, position=position_jitter(), color='blue')
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
sample.ids <- sample( levels(yo$id), 16)

ggplot(subset(yo, id %in% sample.ids), aes(x=time, y=price)) +
  facet_wrap(~id) + 
  geom_line() + 
  geom_point(aes(size=all.purchases), pch=1)
```

```{r}
set.seed(1)
sample.ids <- sample( levels(yo$id), 16)

ggplot(subset(yo, id %in% sample.ids), aes(x=time, y=price)) +
  facet_wrap(~id) + 
  geom_line() + 
  geom_point(aes(size=all.purchases), pch=1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
set.seed(1)
sample.ids <- sample( levels(yo$id), 16)

ggplot(subset(yo, id %in% sample.ids), aes(x=time, y=price)) +
  facet_wrap(~id) + 
  geom_line() + 
  geom_point(aes(size=all.purchases), pch=1)
```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

