---
title: "Lesson 3 - Exploring One Variable"
output: html_notebook
---

## Reading in Data
```{r}
getwd()
list.files()
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)
```

## Histogram of Users' Birthdays
```{r}
install.packages('ggplot2')
library(ggplot2)

1
```

# Use ggplot instead of qplot.
# Introduce bin_width and scale_x_continuous to display data for each day
```{r}
ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31)
```
## Faceting
# Create one graph for each month.
# Use facet_wrap(~variable), break data using category variable
```{r}
ggplot(data = pf, aes(x = dob_day)) +
  geom_histogram(binwidth = 31) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month)
```
## Friend Count
```{r}
ggplot(data = pf, aes(x = friend_count)) +
  geom_histogram()
```
## Limiting the Axes
```{r}
ggplot(data = pf, aes(x = friend_count)) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000))
```
## Adjusting the bin width and add breaks to add step
```{r}
ggplot(data = pf, aes(x = friend_count)) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```

# Facet the histogram for friend counts by gender
```{r}
ggplot(data = pf, aes(x = friend_count)) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

# Omitting NA values
# Subset the data using data = subset() or na.omit
```{r}
ggplot(data = pf, aes(x = friend_count), subset(pf, !is.na(gender))) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```
## Statistics 'by' gender
# 'by' command takes 3 inputs = a variable, a categorical variable, function
# get summary of friend count by gender
```{r}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

## Tenure measured in dats
```{r}
ggplot(aes(x = tenure), data = pf) +
  geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')
```
# Tenure measured in years
```{r}
ggplot(aes(x = tenure/365), data = pf) +
  geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')
```
# Modify x-axis so that it increments in one year
```{r}
ggplot(aes(x = tenure/365), data = pf) +
  geom_histogram(binwidth = .25, color = 'black', fill = '#F79420') +
  scale_x_continuous(limits = c(0, 7), breaks = seq(1, 7, 1))
```
## Labeling Plots
```{r}
ggplot(aes(x = tenure / 365), data = pf) +
  geom_histogram(color = 'black', fill = '#F79420') +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')
```
## Histogram of Facebook Users based on Age
```{r}
ggplot(aes(x = age), data = pf) +
  geom_histogram(binwidth = 1, fill = '#5760AB') +
  scale_x_continuous(breaks = seq(0, 113, 5))
```
## Installing gridExtra
```{r}
install.packages('gridExtra')
library(gridExtra)
```

## Transforming Data
```{r}
summary(pf$friend_count)

# Scale to log10
summary(log10(pf$friend_count))

# To remove negative/-infinity data, add plus one
summary((log10(pf$friend_count+1)))
```
## Create Multiple Plots in One Image Output
```{r}
# can use scale_x_log10 and scale_x_sqrt
library(ggplot2)
p1<- ggplot(data = pf, aes(x = friend_count)) + geom_histogram()
p2 <- ggplot(data = pf, aes(x = log10(friend_count))) + geom_histogram()
p3 <- ggplot(data = pf, aes(x = sqrt(friend_count))) + geom_histogram()
library(gridExtra)
grid.arrange(p1, p2, p3, ncol=1, top="Creating Multiple Plots in One Output")
```
## Add a scaling layer
## Observe the difference in x-axis label
```{r}
p1<- ggplot(data = pf, aes(x = friend_count)) + geom_histogram() 
p2 <- ggplot(data = pf, aes(x = friend_count)) + geom_histogram() +scale_x_log10()
p3 <- ggplot(data = pf, aes(x = friend_count)) + geom_histogram() + scale_x_sqrt()
grid.arrange(p1, p2, p3, ncol=1, top="Creating Multiple Plots in One Output")
```
## Frequency Polygon
## The shape of the frequency polygon depends on how our bins are set up - the height of the lines are the same as the bars in individual histograms, but the lines are easier to make a comparison with since they are on the same axis.
```{r}
ggplot(aes(x = friend_count, y = ..count../sum(..count..)),
       data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth=10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Proportion of users with that friend count')
```
# Likes on the Web
## Note that sum(..count..) will sum across color, so the proportions displayed are based on total users. To plot these proportions within each group, you can try y = ..density...
## Plot for which gender makes more likes
```{r}
ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_log10()
```
```{r}
by(pf$www_likes, pf$gender, sum)
```
## Box Plots, Quartiles, And Friendships
```{r}
ggplot(aes(x = www_likes, y = friend_count), data = subset(pf, !is.na(gender))) +
  geom_boxplot(aes(color = gender), outlier.shape = 1) +
  coord_cartesian(ylim = c(0,1000))
  
```
```{r}
ggplot(aes(x = www_likes, y = friend_count), data = subset(pf, !is.na(gender))) +
  geom_boxplot(aes(color = gender), outlier.shape = 1) +
  coord_cartesian(ylim = c(0,250))
```
```{r}
by(pf$friend_count, pf$gender, summary)
```
```{r}
by(pf$friendships_initiated, pf$gender, summary)
```
## Getting Logical
```{r}
summary(pf$mobile_likes)

summary(pf$mobile_likes > 0)

mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

```
## % check in using mobile
## The sum() function will not work since mobile_check_in is a factor variable. You can use the length() function to determine the number of values in a vector.
```{r}
sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```

