---
title: "Lesson 4 - Problem Set - Exploring One Variable"
output: html_notebook
---

1. Load the Diamonds dataset
```{r}
library(ggplot2)
data(diamonds)
summary(diamonds)
?diamonds
View(diamonds)
str(diamonds)

```

2. Create a histogram of the price of all the diamonds in the diamond data set.
```{r}
ggplot(diamonds, aes(x=price)) +
  geom_histogram(color="black", fill="orange", binwidth= 1000) +
  scale_x_continuous( breaks=seq(0,20000,2000)) +
  xlab("Price") + ylab("Count")
```
3. Diamonds Counts
Supposed we want to know the following:

  - How many cost less than U$500?
  - How many cost less than U$250?
  - How many cost equal to U$15,000 or more?
```{r}
sum(diamonds$price < 500)
sum(diamonds$price < 250)
sum(diamonds$price >= 15000)
```
4. Explore the highest peak 
```{r}
ggplot(diamonds, aes(x=price)) +
  geom_histogram(color="black", fill="orange", binwidth= 50) +
  scale_x_continuous( breaks=seq(0,2000,100)) +
  xlab("Price") + ylab("Count")
```
Lets explore the highest peak further by limiting the values on  x-axis
```{r}
ggplot(diamonds, aes(x = price)) + 
  geom_histogram(color = "black", fill = "DarkOrange", binwidth = 100) + 
  ggtitle("Diamond Price Distribution") +
  scale_x_continuous(breaks = seq(0, 2500, 500)) + 
  coord_cartesian(c(0,2500)) +
  xlab("Price") + ylab("Count" + theme_minimal()) 
```
5. Price by Cut histograms
```{r}
ggplot(data=diamonds) + 
  geom_histogram(binwidth=100, aes(x=diamonds$price)) + 
  ggtitle("Diamond Price Distribution by Cut") + 
  xlab("Diamond Price U$") + ylab("Frequency") + 
  theme_minimal() + 
  facet_wrap(~cut)
```
  - Which cut has the highest priced diamond?
  
```{r}
# by(diamonds$price, diamonds$cut, summary)
subset(diamonds, price == max(price))
```
  - Which cut has the lowest priced diamonds?
  
  
```{r}
subset(diamonds, price == min(price))
```
  - Which cut has the lowest median price?
```{r}
a = diamonds[which(diamonds$cut == 'Fair'), ]
b = diamonds[which(diamonds$cut == 'Good'), ]
c = diamonds[which(diamonds$cut == 'Very Good'), ]
d = diamonds[which(diamonds$cut == 'Premium'), ]
e = diamonds[which(diamonds$cut == 'Ideal'), ]

median(a$price)
median(b$price)
median(c$price)
median(d$price)
median(e$price)
```

6. Scales and Multiple Histograms.
For each cut, the median and quartiles are reasonably close to each other.
But our grid histogram doesnt show it that way. 
The 'Fair' and 'Good' diamonds appear to have different distributions compared to the better cut diamonds. They seem somewhat uniform on the left with long tails on the right.

To make the graph look more similar, lets get different frequency scales(y axis) to accomodate for each cut.
```{r}
ggplot(data=diamonds) + 
  geom_histogram(binwidth=100, aes(x=diamonds$price)) + 
  ggtitle("Diamond Price Distribution by Cut") + 
  xlab("Diamond Price U$") + ylab("Frequency") + 
  theme_minimal() + 
  facet_wrap(~cut, scales = "free_y")
```
We can now observe how different graphs have different Y scales. For example, Ideal cut diamonds have a Y scale maximizing at 200, while Good diamonds have a Y scale max at 400. This is achieved using  scale="free_y" in the facet_wrap layer.


7. Price per Carat by Cut

Create histogram of price per carat and facet it by cut. Adjust the bin width and transform the scale of the x-axis using log10.

```{r}
ggplot(data=diamonds) + 
  geom_histogram(aes(x=diamonds$price/diamonds$carat)) +
  ggtitle("Diamond Price per Carat Distribution by Cut") + 
  xlab("Diamond Price per Carat U$") + 
  ylab("Frequency") + 
  theme_minimal() +
  facet_wrap(~cut, scales="free") + 
  scale_x_log10()
```
8. Price Box Plots
```{r}
# Investigate the price of diamonds using box plots,
# numerical summaries, and one of the following categorical
# variables: cut, clarity, or color.
ggplot(diamonds, aes(factor(cut), price, fill=cut)) + 
  geom_boxplot() + 
  ggtitle("Diamond Price according Cut") + 
  xlab("Type of Cut") + ylab("Diamond Price U$") + 
  coord_cartesian(ylim=c(0,7500))

ggplot(diamonds, aes(factor(clarity), price, fill=clarity)) + 
  geom_boxplot() + 
  ggtitle("Diamond Price according Clarity") + 
  xlab("Clarity") + ylab("Diamond Price U$") + 
  coord_cartesian(ylim=c(0,7500))

ggplot(diamonds, aes(factor(color), price, fill=color)) + 
  geom_boxplot() + 
  ggtitle("Diamond Price according Color") + 
  xlab("Color") + ylab("Diamond Price U$") + 
  coord_cartesian(ylim=c(0,7500))

```

9. Interquartile Range
  - What is the IQR for diamonds with the best color?
  - What is the IQR for diamonds with the worst color?
```{r}
# IQR for the best color
IQR(diamonds$price[diamonds$color == 'D'])


# IQR for the WORST color
IQR(diamonds$price[diamonds$color == 'J'])
```
10. Price Per Carat Box Plots By Color
```{r}
ggplot(diamonds, aes(factor(color), (price/carat), fill=color)) + 
  geom_boxplot() + 
  ggtitle("Diamond Price per Carat according Color") + 
  xlab("Color") + 
  ylab("Diamond Price per Carat U$")
```
11. Carat Frequency Polygon
```{r}
ggplot(data=diamonds, aes(x=carat)) + 
  geom_freqpoly(binwidth = 0.025) + 
  ggtitle("Diamond Frequency by Carat") + 
  xlab("Carat Size") + 
  ylab("Count")+ 
  scale_x_continuous(minor_breaks = seq(0, 5.5, 0.1))

qplot(data=diamonds, x=carat, xlab='Carat', ylab='Frequency', binwidth=0.1, geom='freqpoly') + scale_x_continuous(breaks=seq(0,5,0.2)) + scale_y_continuous(breaks=seq(0,12000,2000))
```
12. Gapminder Data
```{r}
# The Gapminder website contains over 500 data sets with information about
# the world's population. Your task is to download a data set of your choice
# and create 2-5 plots that make use of the techniques from Lesson 3.

# You might use a simple histogram, a boxplot split over a categorical variable,
# or a frequency polygon. The choice is yours!

# You can find a link to the Gapminder website in the Instructor Notes.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. any questions you answered, your observations, and summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.
# ====================================================================================
library('xlsx', quietly=TRUE)
library(reshape2)

sugar <- read.xlsx('indicator_sugar_consumption.xlsx', sheetIndex=1)

# get rid of NA rows and columns
sugar[ncol(sugar)] <- NULL
sugar <- sugar[!is.na(sugar[[1]]),]

# clean up the country column
colnames(sugar)[1] <- 'country'
sugar$country <- gsub(' ','_',sugar$country)

# the countries and years
countries <- as.character(sugar$country)
years <- seq(1961,2004)


# sugar consumption in 2004 histogram
ggplot(sugar, aes(x=X2004)) + geom_histogram() + xlab('Sugar per Person (g per day)') + ylab('Count') + ggtitle('Sugar Consumption per Person by Country in 2004')


# sugar consumption in 1970 vs. 1980 vs. 1990 vs. 2000 boxplot
sugar2 <- melt(sugar, id.vars='country', measure.vars=c('X1970','X1980','X1990','X2000'), na.rm=TRUE, variable.name='year')
sugar2$year <- gsub('X','',sugar2$year)
ggplot(sugar2, aes(x=year, y=value)) + geom_boxplot() + xlab('Year') + ylab('Sugar per Person (g per day)') + ggtitle('Sugar Consumption Worldwide by Decade')
```
13. Explore your friends birthday
```{r}
# Your task is to investigate the distribution of your friends'
# birth months and days.

# Here some questions you could answer, and we hope you think of others.

# **********************************************************************

# How many people share your birthday? Do you know them?
# (Reserve time with them or save money to buy them a gift!)

# Which month contains the most number of birthdays?

# How many birthdays are in each month?

# Which day of the year has the most number of birthdays?

# Do you have at least 365 friends that have birthdays on everyday
# of the year?

# **********************************************************************

# You will need to do some data munging and additional research to
# complete this task. This task won't be easy, and you may encounter some
# unexpected challenges along the way. We hope you learn a lot from it though.

# You can expect to spend 30 min or more on this task depending if you
# use the provided data or obtain your personal data. We also encourage you
# to use the lubridate package for working with dates. Read over the documentation
# in RStudio and search for examples online if you need help.

# You'll need to export your Facebooks friends' birthdays to a csv file.
# You may need to create a calendar of your Facebook friends' birthdays
# in a program like Outlook or Gmail and then export the calendar as a
# csv file.

# Once you load the data into R Studio, you can use the strptime() function
# to extract the birth months and birth days. We recommend looking up the
# documentation for the function and finding examples online.

# We've included some links in the Instructor Notes to help get you started.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. any questions you answered, your observations, and summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation below the line. Submit it when you are ready.
# ===============================================================================
library(lubridate)
library(scales)
bdays <- read.csv('birthdaysExample.csv')
bdays$dates <- as.Date(bdays$dates, '%m/%d/%y')

# to simplify working with February 29th birthdays, set the year to be a leap year
year(bdays$dates) <- 2016

ggplot(bdays, aes(x=dates)) + 
  geom_histogram(binwidth = 1, color = "gray", fill = "blue") + 
  scale_x_date(labels = date_format('%b'), breaks = date_breaks("months"), limits=c(as.Date("2016-01-01"), as.Date("2016-12-31"))) + 
  xlab('Birthday') + ylab('Count') + ggtitle('Histogram of Birthdays')

ggplot(bdays, aes(x=month(bdays$date))) + geom_bar() + scale_x_continuous(breaks=seq(1,12), labels=month.abb) + 
  xlab('Month') + ylab('Number of Birthdays') + ggtitle('Birthdays by Month')
```

